li---
title: "Springs Junction Rainfall"
subtitle: "Data from the automatic rain gauge"
author: "Peter Alspach"
date: "January 2018"
output:
  html_document:
  # html_notebook:
  theme: spacelab
toc: yes
toc_depth: 5
toc_float: yes
code_folding: hide
---
  
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo=TRUE, cache=TRUE, error=TRUE)
knitr::opts_chunk$set(echo=TRUE, results=FALSE, cache=FALSE, error=TRUE)
```

## Get the data
The data is first obtained from the [West Coast Regional Council](http://data.wcrc.govt.nz/cgi-bin/hydwebserver.cgi/points/samples/export?point=597), and saved in C:/Users/PeterAlspach/downloads/github/SpringsJunction/Rainfalldepth.csv. Choose the desired date range, but leave the Format and Aggregate as 'Default'.  This can then be read into R, and manipulated a little.

```{r getData}
rainfall <- read.csv('Rainfall_depth.csv', header=FALSE, skip=1)
names(rainfall) <- c('dateTime','rain')
rainfall$date <- as.Date(substring(rainfall[,1], 1, 8), format='%d/%m/%y')
rainfall <- rainfall[nrow(rainfall):1,]
rainfall$time <- substring(rainfall[,1], 10)
```

```{r findRuns}
# It is extremely unlikely that the 0.5 cup will tip twice in a minute.
# This code checks for 
# such occurrences
tt <- rle(as.character(rainfall$dateTime))$lengths
rainfall$toChk <- rep(tt, tt)
```

## Aggregate the data
The raw data is aggregated daily, as is the data after removing points where the gauge tipped more than once in a minute^[Note, the checking could be improved so that if there is only one tip in a minute, but this is immediately preceded or succeeded by instances of two or more tips per minute these data are also removed.].

```{r aggregate}
aggRain <- aggregate(rainfall$rain, list(rainfall$date), sum)
mthRain <- aggregate(aggRain[,2], list(substring(aggRain[,1], 1, 7)), sum)
lmt <- 1 # maximum number of tips per minute
aggCorr <- aggregate(rainfall[rainfall$toChk <= lmt,'rain'], 
                     list(rainfall[rainfall$toChk <= lmt,'date']), sum)
mthCorr <- aggregate(aggCorr[,2], list(substring(aggCorr[,1], 1, 7)), sum)
mthBrks <- cumsum(rle(substring(aggRain[,1], 1, 7))$lengths)
```

## Plot the rainfall
Then the daily rainfall plots can be produced.  First, the raw data and then the 'corrected' data.  Monthly totals are shown along the top of each plot.
```{r plotRaw, fig.height=8}
par(mfrow=c(2,1), mar=c(4,3,0.5,0.5), mgp=c(2,0.8,0))
with(aggRain, barplot(x, ylab='Daily rainfall (mm)', space=0))
axis(2, c(0, max(aggRain[,2])), c('',''), tck=0)
axis(1, mthBrks[-length(mthBrks)], rep('', length(mthBrks)-1), pos=0) 
text((mthBrks+c(0, mthBrks[-length(mthBrks)]))/2,
     par()$usr[3]-diff(par()$usr[3:4])/20,
     month.name[as.numeric(substring(mthRain[,1], 6))],
     xpd=TRUE, cex=0.9)
text((mthBrks+c(0, mthBrks[-length(mthBrks)]))/2,
     par()$usr[4], round(mthRain[,2]), xpd=TRUE, cex=0.9)
segments(mthBrks, 0, mthBrks, par()$usr[4], lty='dotted')

with(aggCorr, barplot(x, ylab='Daily rainfall (mm)', space=0))
axis(2, c(0, max(aggCorr[,2])), c('',''), tck=0)
axis(1, mthBrks[-length(mthBrks)], rep('', length(mthBrks)-1), pos=0) 
text((mthBrks+c(0, mthBrks[-length(mthBrks)]))/2,
     par()$usr[3]-diff(par()$usr[3:4])/20,
     month.name[as.numeric(substring(mthRain[,1], 6))],
     xpd=TRUE, cex=0.9)
text((mthBrks+c(0, mthBrks[-length(mthBrks)]))/2,
     par()$usr[4], round(mthCorr[,2]), xpd=TRUE, cex=0.9)
segments(mthBrks, 0, mthBrks, par()$usr[4], lty='dotted')
```
