---
title: "Maruia Trap-lines: Trap Locations"
author: "Peter Alspach"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    pdf_document:
  # html_document:
  # html_notebook:
    theme: spacelab
    toc: yes
    toc_depth: 5
    # toc_float: yes
    # code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=FALSE, error=TRUE)
library(knitr)
library(maptools)
```
This code reads the trap locations (longitudes and latitudes) from the GPS waypoint files, and saves them as an R dataset for later use.

It is now deprecated.  To update the trap coordinates it is easiest to load the existing file (load('TrapCoordinates.RData')) and modify it.  To load a new coordinates file: save the GPX file, use an online converter to change it to a csv file, edit in LibreOffice, then read into R.

```{r readWayPts, include=FALSE}
# Read the .gpx files
trapLoc01 <- readGPS('gpx', 'Alfred4x4.gpx', 'w')
trapLoc02 <- readGPS('gpx', 'Campground.gpx', 'w')
trapLoc03 <- readGPS('gpx', 'FergusonsFlat.gpx', 'w')
trapLoc04 <- readGPS('gpx', 'Kellys.gpx', 'w')
trapLoc05 <- readGPS('gpx', 'LakeDaniellTerrace.gpx', 'w')
trapLoc06 <- readGPS('gpx', 'LakeDaniellTrack.gpx', 'w')
trapLoc07 <- readGPS('gpx', 'Lockingtons.gpx', 'w')
trapLoc08 <- readGPS('gpx', 'MarbleHill.gpx', 'w')
trapLoc09 <- readGPS('gpx', 'MaruiaRiver.gpx', 'w')
trapLoc10 <- readGPS('gpx', 'PellTerrace.gpx', 'w')
trapLoc11 <- readGPS('gpx', 'Reids.gpx', 'w')
trapLoc12 <- readGPS('gpx', 'StateHighway.gpx', 'w')
```

The reading isn't perfect and thus each dataframe needs a bit of manual adjustment.

```{r correctDF, include=FALSE}
# Alfred4x4
tt01 <- trapLoc01[, c(2,6,7)]
tt01[,1] <- as.character(tt01[,1])
for (i in 2:3) tt01[,i] <- as.numeric(as.character(tt01[,i]))
names(tt01) <- c('Trap','S','E')
for (i in 6:13) trapLoc01[,i] <- as.numeric(as.character(trapLoc01[,i]))
tt01[c(39,46), 2:3] <- trapLoc01[c(39,46), 10:11]
tt01 <- tt01[!is.na(tt01[,2]),]
tt01 <- tt01[order(tt01[,1]),]

# Campground (no adjustment necessary)
tt02 <- trapLoc02[, c(2,6,7)]
tt02[,1] <- as.character(tt02[,1])
for (i in 2:3) tt02[,i] <- as.numeric(as.character(tt02[,i]))
names(tt02) <- c('Trap','S','E')

# FergusonsFlat
tt03 <- trapLoc03[, c(2,6,7)]
tt03[,1] <- as.character(tt03[,1])
for (i in 2:3) tt03[,i] <- as.numeric(as.character(tt03[,i]))
names(tt03) <- c('Trap','S','E')
for (i in 6:13) trapLoc03[,i] <- as.numeric(as.character(trapLoc03[,i]))
tt03[17, 2:3] <- trapLoc03[17,8:9]
tt03[17,1] <- 'FF24'
tt03 <- tt03[!is.na(tt03[,2]),]
tt03 <- tt03[order(tt03[,1]),]
plot(tt03[,2], tt03[,3], type='n')
text(tt03[,2], tt03[,3], substring(tt03[,1], 3), cex=0.6)

# Kellys (no adjustment necessary)
tt04 <- trapLoc04[, c(2,6,7)]
tt04[,1] <- as.character(tt04[,1])
for (i in 2:3) tt04[,i] <- as.numeric(as.character(tt04[,i]))
names(tt04) <- c('Trap','S','E')

# LakeDaniellTerrace (no adjustment necessary)
tt05 <- trapLoc05[, c(2,6,7)]
tt05[,1] <- as.character(tt05[,1])
for (i in 2:3) tt05[,i] <- as.numeric(as.character(tt05[,i]))
names(tt05) <- c('Trap','S','E')

# LakeDaniellTrack (no adjustment necessary)
tt06 <- trapLoc06[, c(2,6,7)]
tt06[,1] <- as.character(tt06[,1])
for (i in 2:3) tt06[,i] <- as.numeric(as.character(tt06[,i]))
names(tt06) <- c('Trap','S','E')

# Lockingtons (no adjustment necessary)
tt07 <- trapLoc07[, c(2,6,7)]
tt07[,1] <- as.character(tt07[,1])
for (i in 2:3) tt07[,i] <- as.numeric(as.character(tt07[,i]))
names(tt07) <- c('Trap','S','E')

# MarbleHill (no adjustment necessary)
tt08 <- trapLoc08[, c(2,6,7)]
tt08[,1] <- as.character(tt08[,1])
for (i in 2:3) tt08[,i] <- as.numeric(as.character(tt08[,i]))
names(tt08) <- c('Trap','S','E')

# MaruiaRiver (no adjustment necessary)
tt09 <- trapLoc09[, c(2,6,7)]
tt09[,1] <- as.character(tt09[,1])
for (i in 2:3) tt09[,i] <- as.numeric(as.character(tt09[,i]))
names(tt09) <- c('Trap','S','E')

# PellTerrace
tt10 <- trapLoc10[, c(2,6,7)]
tt10[,1] <- as.character(tt10[,1])
for (i in 2:3) tt10[,i] <- as.numeric(as.character(tt10[,i]))
names(tt10) <- c('Trap','S','E')
for (i in 6:13) trapLoc10[,i] <- as.numeric(as.character(trapLoc10[,i]))
tt10[30, 2:3] <- trapLoc10[38,7:8]
tt10 <- tt10[!is.na(tt10[,2]),]
tt10 <- tt10[order(tt10[,1]),]

# Reids
tt11 <- trapLoc11[, c(2,6,7)]
tt11[,1] <- as.character(tt11[,1])
for (i in 2:3) tt11[,i] <- as.numeric(as.character(tt11[,i]))
names(tt11) <- c('Trap','S','E')
for (i in 6:13) trapLoc11[,i] <- as.numeric(as.character(trapLoc11[,i]))
tt11[c(23,27), 2:3] <- trapLoc11[c(23,27), 10:11]
tt11 <- tt11[!is.na(tt11[,2]),]
tt11 <- tt11[order(tt11[,1]),]

# StateHighway (no adjustment necessary)
tt12 <- trapLoc12[, c(2,6,7)]
tt12[,1] <- as.character(tt12[,1])
for (i in 2:3) tt12[,i] <- as.numeric(as.character(tt12[,i]))
names(tt12) <- c('Trap','S','E')
```

```{r combineAll}
trapLoc <- rbind(tt01, tt02, tt03, tt04, tt05, tt06, tt07, tt08, tt09, tt10, tt11, tt12)
trapLoc <- trapLoc[order(trapLoc[,1]),]
save(trapLoc, file='TrapCoordinates.RData')
```
