---
title: "Maruia Trap-lines: Monthly Summary"
author: "Peter Alspach"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    pdf_document:
  # html_document:
  # html_notebook:
    theme: spacelab
toc: yes
toc_depth: 5
    # toc_float: yes
    # code_folding: hide
---

<!-- To do: -->
<!-- Mammals are defined at the end of the 'getReady' chunk (about line 45) -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=FALSE, error=TRUE)
library(knitr)
library(kableExtra)
library(jpeg)
library(TraplineTools)

setwd('~/GitHub/SpringsJunction/TrapLines') # set appropriate working directory
```
## Introduction

```{r readData}
# Read the data
tl <- getTrapData(dstOffset=TRUE)
yr <- tl$yr
tl <- tl$tl
```

```{r getReady}
# Set the cut-off dates for the monthly intervals
lastDate <- sort(tl$Date.Checked, decreasing=TRUE)[1]
beginAt <- 10 # Good options are 10/11 and 22/23
obsMth <- setCutDates(beginAt, yr, tl$jDay)
mth <- max(as.numeric(obsMth), na.rm=TRUE)

# Set the standard trap status of interest
stdStat <- c('Sprung and Empty','Rat','Stoat','Weasel') # standard trap status
allMammals <- c('Rat','Stoat','Weasel','Cat','Ferret','Mouse','Rabbit','Hedgehog','Hare','Possum')

# Next compute the number of traps per line (irrespective of whether some are missing)
nTrap <- table(tl$Line, gsub('([[:upper:]]+)([[:digit:]]+)', '\\2', tl$Tag.No))
nTrap <- apply(nTrap, 1, function(x) length(x[x!=0]))
```

```{r readDataPY}
# Read the data for the previous year
tlPY <- getTrapData(paste0('TrapCatches', yr-1, '.csv'), dstOffset=TRUE)$tl

obsMthPY <- setCutDates(beginAt, yr-1, tlPY$jDay)
```

The following tables give a summary of the catches^[Actually, the catch is not of primary interest as the objective of the trapping is to help protect native species (i.e., birds, bats and insects) from predation.  Thus, the desired outcome is an increase in the population of these species.  However, this is difficult to monitor.] from the `r length(unique(tl$Tag.No))` trap boxes^[On some lines (Maruia River, Kellys, and parts of both the Lewis Pass Highway and Fergusons Flat) each box contains two traps.  Thus, on these lines it is possible to catch two mammals per trap box.] on the `r length(unique(tl$Line))` trap-lines at Maruia.  The intention is to check the traps monthly around the change of the month, hence monthly summaries are aggregated from the `r beginAt`^th^ of of the month (inclusive).

The summary is for the most recent monthly interval (i.e., that beginning on `r unique(obsMth[tl$jDay==max(tl$jDay)])`) in `r yr` and also the entire year-to-date.

## The most recent monthly interval

```{r consolidateAll}
# Set up arrays to hold the summaries
catchAllPY <- catchAll <- array(dim=c(length(unique(tl$Line))+1, 1+2*(length(stdStat)+2), 12),
                                dimnames=list(c(unique(tl$Line), 'TOTAL'),
                                              c('nTrap', rep(c(substring(stdStat, 1, 6), 'Other', 'Total'), 2)),
                                              levels(obsMth)))
# Current year
for (i in 1:mth) catchAll[,,i] <- as.matrix(consolidateTrapData(currMth=i)[, 1:dim(catchAll)[2]])

# Previous year
for (i in 1:12) catchAllPY[,,i] <- as.matrix(consolidateTrapData(tlPY, mthInt=obsMthPY, currMth=i)[, 1:dim(catchAll)[2]])
```

```{r currMth}
catch <- catchAll[, , mth]
```

Table \ref{currMthN} (below) summarises the number of mammals caught in the most recent monthly interval for each line.  The first column is the number of trap boxes^[In some cases trap boxes may be missing or not functioning, and these have been omitted.  Hence the numbers here may not correspond to totals elsewhere], followed by the number of traps which were sprung but empty.  The remaining columns (in blue) give the catch of rats, stoats, weasels, other mammals (e.g., cats, mice, rabbits) and the total mammals.  

```{r currMthN}
# kable(catch[, 1:(length(stdStat)+3)]) %>%
#   kable_styling(full_width=FALSE) %>%
kable(catch[, 1:(length(stdStat)+3)], 'latex', booktabs=TRUE,
      caption="\\label{currMthN}Current month's catch for each line. Note: nTrap is the number of functioning traps (i.e., missing and non-functioning traps have been omitted).") %>%
  kable_styling(latex_options="hold_position") %>%
  column_spec(length(stdStat)+4, bold=TRUE, color='blue') %>%
  column_spec(2, color='gray') %>%
  column_spec(4:7, color='blue') %>%
  row_spec(nrow(catch), bold=TRUE, background="#e5e5e5")
```

\pagebreak

The next table (Table \ref{currMthPC}) is similar, except that the numbers are expressed as percentages of the total number of trap boxes^[Omitting missing and not functioning] on the line.

```{r currMthPC}
kable(catch[, c(1, ((length(stdStat)+4):ncol(catch)))], 'latex', booktabs=TRUE,
      caption="\\label{currMthPC}Current month's catch per trap box ($\\times$ 100) for each line.") %>%
  kable_styling(latex_options="hold_position") %>%
  column_spec(length(stdStat)+4, bold=TRUE, color='blue') %>%
  column_spec(2, color='gray') %>%
  column_spec(4:7, color='blue') %>%
  row_spec(nrow(catch), bold = T, background = "#e5e5e5")
```

## The year to date

### Mammals caught by line

```{r yrToDate}
catch <- consolidateTrapData()
```

Table \ref{yrToDateN} (below) summarises the number of mammals caught in the year to date for each line.  The first column is the number of trap boxes, followed by the number of traps which were sprung but empty.  The remaining columns (in blue) give the catch of rats, stoats, weasels, other mammals (e.g., cats, ferrets, rabbits) and the total mammals.  

```{r yrToDateN, echo=FALSE}
kable(catch[, 1:(length(stdStat)+3)], 'latex', booktabs=TRUE,
      caption='\\label{yrToDateN}Total year-to-date catch for each line.') %>%
  kable_styling(latex_options="hold_position") %>%
  column_spec(length(stdStat)+4, bold=TRUE, color='blue') %>%
  column_spec(2, color='gray') %>%
  column_spec(4:7, color='blue') %>%
  row_spec(nrow(catch), bold = T, background = "#e5e5e5")
```

\pagebreak

Next, the data are presented as numbers caught per trap box per month ($\times$ 100) for each line (Table \ref{yrToDatePC}).  This is analogous to the second table in the current month section which gave the percentages.

```{r yrToDatePC}
kable(catch[, c(1, ((length(stdStat)+4):ncol(catch)))], 'latex', booktabs=TRUE,
      caption='\\label{yrToDatePC}Year-to-date catch per trap box per month ($\\times$ 100) for each line.') %>%
  kable_styling(latex_options="hold_position") %>%
  column_spec(length(stdStat)+4, bold=TRUE, color='blue') %>%
  column_spec(2, color='gray') %>%
  column_spec(4:7, color='blue') %>%
  row_spec(nrow(catch), bold = T, background = "#e5e5e5")
```

\pagebreak

### Mammals caught by month

```{r mthCatchP}
# Current year
catch <- t(catchAll['TOTAL',,])

# Previous year
catchPY <- t(catchAllPY['TOTAL',,])
```

Figure \ref{mthCatchPFig} shows the catch per month summed over all lines for all mammals (total), rats, mustelids (stoats and weasels) and other mammals (including ferrets).  The figure includes the previous year's results for comparison.

```{r, mthCatchPFig, fig.height=4, fig.cap="\\label{mthCatchPFig}Trap catches summed over all lines for each monthly interval ('Sprung' means the trap was tripped but empty). Previous year represented by dashed lines with the same colour coding.  Note: 'mustelid' includes only stoats and weasels (ferrets, if present, are included in 'other').  Circle diameter, shown for the Total catch only but applicable to all categories, is proportional to the mean number of times the traps were checked during the month"}
par(mar=c(3,3,0.5,2), las=1, mgp=c(2.2, 0.8, 0), xpd=TRUE)
tt <- catch[!(catch[,'Total']==0 | is.na(catch[,'Total'])),]
ttPY <- catchPY
ttPY[ttPY[,'Total']==0,] <- NA
yLim <- c(0, max(c(tt[,'Total'], ttPY[,'Total']), na.rm=TRUE))
plot(1:nrow(tt), tt[,'Total'], xlab='Month beginning', ylab='Numbers caught',
     xlim=c(1,12), ylim=yLim, xaxt='n', type='l', bty='n')
axis(1, 1:nrow(ttPY), row.names(ttPY))
axis(2, yLim, c('',''), tck=0)            
points(1:nrow(tt), tt[,'Total'], pch=16, cex=table(obsMth)/(2*length(unique(tl$Tag.No))))
lines(1:nrow(tt), tt[,'Rat'], col='gray40')
lines(1:nrow(tt), tt[,'Stoat']+tt[,'Weasel'], col='brown')
lines(1:nrow(tt), tt[,'Other'], col='orange')
lines(1:nrow(tt), tt[,'Sprung'], col='green')

# Previous year
points(1:nrow(ttPY), ttPY[,'Total'], pch=1, cex=table(obsMthPY)/(2*length(unique(tlPY$Tag.No))))
lines(1:nrow(ttPY), ttPY[,'Total'], lty='dashed')
lines(1:nrow(ttPY), ttPY[,'Rat'], col='gray40', lty='dashed')
lines(1:nrow(ttPY), ttPY[,'Stoat']+ttPY[,'Weasel'], col='brown', lty='dashed')
lines(1:nrow(ttPY), ttPY[,'Other'], col='orange', lty='dashed')
lines(1:nrow(ttPY), ttPY[,'Sprung'], col='green', lty='dashed')

text(rep(nrow(tt)+0.1, 5), c(tt[nrow(tt),c('Total','Rat')]+c(2,-2),
                             tt[nrow(tt),'Stoat']+tt[nrow(tt),'Weasel']+2, tt[nrow(tt),'Other']-2,
                             tt[nrow(tt),'Sprung']),
     c('Total','Rat','Mustelid','Other','Sprung'), adj=0, xpd=TRUE, cex=0.8,
     col=c('black','gray40','brown','orange','green'))
```

\pagebreak

### Catch by month for each line

```{r mthCatchL}
# Current year
catch <- t(catchAll[,dim(catchAll)[2],])[1:mth,]
catch <- catch[, -ncol(catch)] # remove TOTAL column

# Previous year
catchPY <- t(catchAllPY[,dim(catchAll)[2],])
catchPY <- catchPY[, -ncol(catchPY)] # remove TOTAL column
```

Figure \ref{mthCatchLFig} shows the monthly catch per trap box ($\times$ 100) of all mammals for half of the lines on the top and for the other half on the bottom.  The colours have been chosen to roughly match those on the map supplied by DoC, and the line names have been abbreviated.  The previous year's results are also shown for comparison (although it is not always easy to distinguish the colours).

```{r, mthCatchLFig, fig.height=6.3, fig.cap="\\label{mthCatchLFig}Monthly trap catch per trap ($\\times$ 100) of all mammals for each line. Some lines were not checked in some monthly periods, which is indicated by the disconnected lines. Previous year represented by dashed lines with the same colour coding. Note: the vertical scale on the two figures may differ."}
par(mar=c(3,3,0,2.5), las=1, mgp=c(2.2, 0.8, 0), mfrow=c(2,1))
tt <- catch[, colnames(catch) %in% c('Lewis Pass H/Way','Lockingtons','Marble Hill',
                                     'Maruia River','Kellys','Campground')]
ttPY <- catchPY[, colnames(catchPY) %in% c('Lewis Pass H/Way','Lockingtons','Marble Hill',
                                     'Maruia River','Kellys','Campground')]
lCol <- c('gold','darkgreen','grey40','turquoise','lightgreen','pink3')
yLim <- c(0, max(c(tt, ttPY), na.rm=TRUE))
plot(1:nrow(tt), tt[,1], xlab='Month beginning', ylab='Catch per trap (x100)',
     xlim=c(1,12), ylim=yLim, xaxt='n', type='l', bty='n')
axis(1, 1:12, levels(obsMth))
axis(2, yLim, c('',''), tck=0)

for(i in 1:ncol(tt))
{
  lines(1:nrow(ttPY), ttPY[,i], col=lCol[i], lty='dashed')
  points(1:nrow(ttPY), ttPY[,i], col=lCol[i], cex=0.5)
  lines(1:nrow(tt), tt[,i], col=lCol[i])
  points(1:nrow(tt), tt[,i], col=lCol[i], pch=16, cex=0.5)
  text(nrow(tt)+0.1, tt[nrow(tt), i], abbreviate(colnames(tt), 6)[i],
       adj=0, xpd=TRUE, cex=0.8, col=lCol[i])
}

tt <- catch[, !(colnames(catch) %in% c('Lewis Pass H/Way','Lockingtons','Marble Hill',
                                       'Maruia River','Kellys','Campground'))]
ttPY <- catchPY[, !(colnames(catchPY) %in% c('Lewis Pass H/Way','Lockingtons','Marble Hill',
                                       'Maruia River','Kellys','Campground'))]
lCol <- c('darkred','pink3','grey40','blue','red','green')
yLim <- c(0, max(c(tt, ttPY), na.rm=TRUE))
plot(1:nrow(tt), tt[,1], xlab='Month beginning', ylab='Catch per trap (x100)',
     xlim=c(1,12), ylim=yLim, xaxt='n', type='l', bty='n')
axis(1, 1:12, levels(obsMth))
axis(2, yLim, c('',''), tck=0)

for(i in 1:ncol(tt))
{
  lines(1:nrow(ttPY), ttPY[,i], col=lCol[i], lty='dashed')
  points(1:nrow(ttPY), ttPY[,i], col=lCol[i], cex=0.5)
  lines(1:nrow(tt), tt[,i], col=lCol[i])
  points(1:nrow(tt), tt[,i], col=lCol[i], pch=16, cex=0.5)
  text(nrow(tt)+0.1, tt[nrow(tt), i], abbreviate(colnames(tt), 6)[i],
       adj=0, xpd=TRUE, cex=0.8, col=lCol[i])
}
```

## Miscellaneous species

```{r, miscSpp}
miscSppY <- table(unlist(tl[, c('Trap.1','Trap.2')]))
miscSppM <- table(unlist(tl[obsMth==levels(obsMth)[mth], c('Trap.1','Trap.2')]))
```

### Other mammals

The other mammals (i.e., 'Other' in the tables and figures above) caught so far this year are:
`r mSppFn(miscSppY, 'Cat', 'cats')`,
`r mSppFn(miscSppY, 'Ferret', 'ferrets')`,
`r mSppFn(miscSppY, 'Hedgehog', 'hedgehogs')`,
`r mSppFn(miscSppY, 'Mouse', 'mice')`, and
`r mSppFn(miscSppY, 'Rabbit', 'rabbits')`.
Of these, `r if ('Cat' %in% names (miscSppM)) paste0(mSppFn(miscSppM, 'Cat', 'cats', zero=FALSE), ',')` `r if ('Ferret' %in% names (miscSppM)) paste0(mSppFn(miscSppM, 'Ferret', 'ferrets', zero=FALSE), ',')` `r if ('Hedgehog' %in% names (miscSppM)) paste0(mSppFn(miscSppM, 'Hedgehog', 'hedgehogs', zero=FALSE), ',')` `r if ('Mouse' %in% names (miscSppM)) paste0(mSppFn(miscSppM, 'Mouse', 'mice', zero=FALSE), ',')` `r if ('Rabbit' %in% names (miscSppM)) paste0(mSppFn(miscSppM, 'Rabbit', 'rabbits', zero=FALSE), ',')` was/were caught in the current monthly period.

### Non-target species

Finally, in the year to date, the following non-target species have been caught: 
`r mSppFn(miscSppY, 'Harrier', 'harrier')`,
`r mSppFn(miscSppY, 'Robin', 'robins')`, 
`r mSppFn(miscSppY, 'Weka', 'weka')`, and
`r mSppFn(miscSppY, 'Weka Juvenile', 'weka juveniles')`.

```{r OtherSeeComment}
osc <- tl[tl$Trap.1=='Other see comment' | tl$Trap.2=='Other see comment',]
osc <- osc[osc$Record.Comment!='' & osc$Status=='Open',]
```
`r if(nrow(osc)>0) paste('Other non-target species caught in the year to date are', osc$Record.Comment)`.
\color{black!50}

### Check

Here are the contents of \texttt{miscSppY} and \texttt{miscSppM} as a check that nothing has been missed from the two previous sub-sections:

```{r miscSppChk}
chkTab <- merge(cbind(Spp=names(miscSppY), ToDate=miscSppY),
                cbind(Spp=names(miscSppM), Month=miscSppM), all=TRUE)
excl <- c('N/A','Other see comment','Rat','Sprung and Empty','Still Set',
          'Still Set Bait Eaten','Stoat','Weasel')
chkTab <- chkTab[!(chkTab[,'Spp'] %in% excl),]
row.names(chkTab) <- NULL
kable(chkTab, 'latex', booktabs=TRUE) %>%
  kable_styling(latex_options="hold_position") %>%
  column_spec(1:3, color='gray')
```
\pagebreak
\color{black}

## Catch by Trap

It may be of interest to know which traps are catching the most mammals.  This is shown in the following 'maps'^[No attempt has been made to ensure that the aspect ratio of the maps is correct.  This allows for the full page to be utilised.] (Figures \ref{mthTrapCatch1}, \ref{mthTrapCatch2} and \ref{mthTrapCatch3}).  These are a little complicated.  Each map shows a collection of lines, and the location of the traps on these lines.  The symbols indicate the cumulative catch over the previous twelve months such that the diameter of the cycle is proportional to the number of mammals caught, and the colour of the ring indicates the number of months prior to the current.  Thus, a lack of yellow-shaded rings indicates that no mammals were caught 9-12 months ago.  Similarly, lack of dark blue or purple shaded rings indicates that no mammals have been caught recently.  Also, one can see traps 4, 9 and 10 on the Campground have not caught anything during the last year, which is because these traps are missing.

```{r mthCatchT}
# Current year
catch <- catchPerBox()

# Previous year
catchPY <- catchPerBox(tlPY, obsMthPY)

# Combine catch and catchPY to get 12 months data in order
catch <- t(catch)
catchPY <- t(catchPY)
catch <- catch[, apply(catch, 2, function(x) length(x[x!=0 & !is.na(x)])!=0)]
catch12m <- cbind(catchPY[,-(1:ncol(catch))], catch)
```

```{r trapCoords}
load('TrapCoordinates.RData') # Waypoints (saved by TrapLocations.Rmd)

# Rename Tag.No to match Waypoints
trap <- row.names(catch12m)
trap[substring(trap, 1, 1)=='A'] <- sub('A', 'ALF', trap[substring(trap, 1, 1)=='A'])
trap[substring(trap, 1, 1)=='B'] <- sub('B', 'X', trap[substring(trap, 1, 1)=='B']) # dummy substitution
trap[substring(trap, 1, 1)=='E'] <- sub('E', 'EE', trap[substring(trap, 1, 1)=='E'])
trap[substring(trap, 1, 1)=='F'] <- sub('F', 'FF', trap[substring(trap, 1, 1)=='F'])
trap[substring(trap, 1, 1)=='K'] <- sub('K', 'AK', trap[substring(trap, 1, 1)=='K'])
trap[substring(trap, 1, 1)=='M'] <- sub('M', 'MBL', trap[substring(trap, 1, 1)=='M'])
trap[substring(trap, 1, 1)=='X'] <- sub('X', 'MRB', trap[substring(trap, 1, 1)=='X']) # actual substitution
trap[substring(trap, 1, 1)=='P'] <- sub('P', 'PS', trap[substring(trap, 1, 1)=='P'])
trap[substring(trap, 1, 1)=='R'] <- sub('RS', 'RDS', trap[substring(trap, 1, 1)=='R'])
trap[substring(trap, 1, 1)=='S'] <- sub('SH', 'SHY', trap[substring(trap, 1, 1)=='S'])
trap[substring(trap, 1, 1)=='T'] <- sub('T', 'LDT', trap[substring(trap, 1, 1)=='T'])
row.names(catch12m) <- trap

# Merge trapLoc with catch12m
catch12m <- merge(trapLoc, catch12m, by.x='Trap', by.y='row.names')

cumCatch <- catch12m
cumCatch[, ncol(cumCatch):4] <- t(apply(cumCatch[, ncol(cumCatch):4], 1, cumsum))

lineGrps <- vector('list', 3)
lineGrps[[1]] <- c('ALF','LD','LDT','PS','RDS')
lineGrps[[2]] <- c('CG','SHY')
lineGrps[[3]] <- c('AK','CG','EE','FF','MBL','MRB')
```

```{r, mthTrapCatch1, fig.height=10, fig.width=8, fig.cap="\\label{mthTrapCatch1}Map of the Lake Daniell lines (i.e., Lake Daniell Track and Terrace, Alfred 4x4, Pell Terrace and Reid Stream).  Each line is a different colour.  The point diameters are proportional to the cumulative catch, and the colour indicates when the mammal was caught (months prior to the present).  An open grey circle indicates a trap which has caught nothing in the past 12 months.  The topographical map overlay is approximate only.  Note, the GPS locations of traps 46-48 on Reid Stream are not current, and the GPS trap numbers on Alfred 4x4 are slightly out of synchrony with the actual trap numbers."}
par(mar=rep(0, 4))
myMap <- readJPEG('MapFigure3.jpg')
ttCols <- viridisLite::inferno(ncol(catch12m)-3)

# for (lns in 1:length(lineGrps))
lns <- 1
{  
  ttLines <- catch12m[gsub('[[:digit:]]', '', catch12m$Trap) %in% lineGrps[[lns]],]
  ttCum <- cumCatch[gsub('[[:digit:]]', '', cumCatch$Trap) %in% lineGrps[[lns]],]
  with(ttLines, plot(E, S, cex=0.7, col='grey50', axes=FALSE, xlab='', ylab=''))
  ttUsr <- par()$usr
  rasterImage(myMap, ttUsr[1], ttUsr[3]-0.002, ttUsr[2], ttUsr[4]+0.0005)
  with(ttLines, points(E, S, cex=0.7, col='grey50'))
  for (i in 4:ncol(catch12m))
  {  
    points(ttLines[ttLines[,i]!=0, c('E', 'S')], pch=16, col=ttCols[ncol(catch12m)+1-i],
           cex=ttCum[ttLines[,i]!=0, i]/2)
  }
  lCols <- c(ALF='magenta', LD='blue', LDT='darkred', PS='red', RDS='green')
  for (i in lineGrps[[lns]])
  {
    with(ttLines[gsub('[[:digit:]]', '', ttLines$Trap)==i,], lines(E, S, col=lCols[i]))
  }
  legend('bottomright', legend=c('1 catch', '3 catches', '6 catches', '9 catches'),
         pt.cex=c(1, 3, 6, 9)/2, pch=1, pt.lwd=2, bty='n')
  legend('top', legend=c('Current', '3 months ago', '6 months ago', '9 months ago', '12 months ago'),
         pch=16, col=ttCols[c(1,3,6,9,12)], bty='n')
}
# To do:
#   Ensure aspect ratio is correct
#   Add actual map (maybe)
```

```{r, mthTrapCatch2, fig.height=10, fig.width=8, fig.cap="\\label{mthTrapCatch2}Map of the Lewis Pass H/Way and Campground lines.  Each line is a different colour.  The point diameters are proportional to the cumulative catch, and the colour indicates when the mammal was caught (months prior to the present).  An open grey circle indicates a trap which has caught nothing in the past 12 months.  The topographical map overlay is approximate only."}
par(mar=rep(0, 4))
myMap <- readJPEG('MapFigure4.jpg')
ttCols <- viridisLite::inferno(ncol(catch12m)-3)

# for (lns in 1:length(lineGrps))
lns <- 2
{  
  ttLines <- catch12m[gsub('[[:digit:]]', '', catch12m$Trap) %in% lineGrps[[lns]],]
  ttCum <- cumCatch[gsub('[[:digit:]]', '', cumCatch$Trap) %in% lineGrps[[lns]],]
  with(ttLines, plot(E, S, cex=0.7, col='grey50', axes=FALSE, xlab='', ylab=''))
  ttUsr <- par()$usr
  rasterImage(myMap, ttUsr[1], ttUsr[3]-0.0015, ttUsr[2], ttUsr[4]+0.0025)
  with(ttLines, points(E, S, cex=0.7, col='grey50'))
  for (i in 4:ncol(catch12m))
  {  
    points(ttLines[ttLines[,i]!=0, c('E', 'S')], pch=16, col=ttCols[ncol(catch12m)+1-i],
           cex=ttCum[ttLines[,i]!=0, i]/2)
  }
  lCols <- c(CG='gold', SHY='grey40')
  for (i in lineGrps[[lns]])
  {
    with(ttLines[gsub('[[:digit:]]', '', ttLines$Trap)==i,], lines(E, S, col=lCols[i]))
  }
  legend('bottomleft', legend=c('1 catch', '3 catches', '6 catches', '9 catches'),
         pt.cex=c(1, 3, 6, 9)/2, pch=1, pt.lwd=2, bty='n')
  legend('top', legend=c('Current', '3 months ago', '6 months ago', '9 months ago', '12 months ago'),
         pch=16, col=ttCols[c(1,3,6,9,12)], bty='n')
}
# To do:
#   Ensure aspect ratio is correct
#   Add actual map (maybe)
```

```{r, mthTrapCatch3, fig.height=10, fig.width=8, fig.cap="\\label{mthTrapCatch3}Map of the remaining lines (i.e., Campground, Lockingtons, Fergusons Flat, Kellys, Marble Hill and Maruia River).  Each line is a different colour.  The point diameters are proportional to the cumulative catch, and the colour indicates when the mammal was caught (months prior to the present).  An open grey cirlce indicates a trap which has caught nothing in the past 12 months.  The topographical map overlay is approximate only."}
par(mar=rep(0, 4))
myMap <- readJPEG('MapFigure5.jpg')
ttCols <- viridisLite::inferno(ncol(catch12m)-3)

# for (lns in 1:length(lineGrps))
lns <- 3
{  
  ttLines <- catch12m[gsub('[[:digit:]]', '', catch12m$Trap) %in% lineGrps[[lns]],]
  ttCum <- cumCatch[gsub('[[:digit:]]', '', cumCatch$Trap) %in% lineGrps[[lns]],]
  with(ttLines, plot(E, S, cex=0.7, col='grey50', axes=FALSE, xlab='', ylab='', type='n'))
  ttUsr <- par()$usr
  rasterImage(myMap, ttUsr[1]-0.0015, ttUsr[3], ttUsr[2]+0.002, ttUsr[4]-0.004)
  with(ttLines, points(E, S, cex=0.7, col='grey50'))
  for (i in 4:ncol(catch12m))
  {  
    points(ttLines[ttLines[,i]!=0, c('E', 'S')], pch=16, col=ttCols[ncol(catch12m)+1-i],
           cex=ttCum[ttLines[,i]!=0, i]/2)
  }
  lCols <- c(CG='gold', EE='turquoise', FF='pink', AK='darkgreen', MBL='lightgreen', MRB='pink3')
  for (i in lineGrps[[lns]])
  {
    with(ttLines[gsub('[[:digit:]]', '', ttLines$Trap)==i,], lines(E, S, col=lCols[i]))
  }
  legend(ttUsr[2]-0.008, ttUsr[4]-0.008, legend=c('1 catch', '3 catches', '6 catches', '9 catches'),
         pt.cex=c(1, 3, 6, 9)/2, pch=1, pt.lwd=2, bty='n')
  legend('right', legend=c('Current', '3 months ago', '6 months ago', '9 months ago', '12 months ago'),
         pch=16, col=ttCols[c(1,3,6,9,12)], bty='n')
}
# To do:
#   Ensure aspect ratio is correct
#   Add actual map (maybe)
```
